---
title: "R Notebook"
output: html_notebook
---

This [R Markdown](http://rmarkdown.rstudio.com) Notebook contains the code used to produce the analyses of locmotor behavior under temperature cycles in Nematostella vectensis (Berger & Tarrant 2022).

Load libraries and set directory
```{r}
library(behavr)
library(ggetho)
library(damr)
library(scopr)
library(sleepr)
library(zeitgebr)
library(lomb)
library(dplyr)
library(circacompare)
library(pracma)
library(stringr)
library(smooth)
library(data.table)
library(circular)
library(DescTools)
library(ggplot2)
library(car)
library(onewaytests)


setwd("F:/circadian/temperature_cycles/")


```

This next chunk was used to calculate p-values based on the lomb-scargle periodogram (LSP) randomization. Not run because time-consuming. 
```{r, eval=FALSE}

#read in raw data
df = read.csv("master_thin.csv")

#smooths time seires with 4-pt centered moving average
smoothed = apply(df, 1, function(j){
  
  dat = j[-c(1,2)]
  
  NAs = is.na(dat)
  
  smooth = cma(y=as.numeric(t(dat[!(NAs)]) ) , order=4)
  
  dat[!(NAs)] = smooth$fitted
  
  return(dat)
})

colnames(smoothed) = df$ID

write.csv(t(smoothed), file="smoothed_temperature.csv")

##LSP with p-values from 2000 permutations, restricted periods from 20-28
rando_smooth_circ = apply(smoothed, 2, function(j, n=2000){
  tmp = randlsp(repeats = n, x=as.numeric(j), type="period", ofac=50, from = 20, to=28, plot=FALSE, trace=FALSE)
  return(summary(tmp))
})

#also check for circatidal rhythms, periods 10-14
rando_smooth_tidal = apply(smoothed, 2, function(j, n = 2000){
  tmp = randlsp(repeats = n, x=as.numeric(j), type="period", ofac=50, from = 10, to=14, plot=FALSE, trace=FALSE)
  return(summary(tmp))
})


# This function parses the above output and corrects the p-values. P-values from permutation tests should not be zero, so we add 1/nperm to those values and then corect for multiple testing.
p_lsp_fun <- function(x, perm=2000){
  ps = as.numeric(lapply(x, function(i){return(i$Value[13])}))
  power = as.numeric(lapply(x, function(i){return(i$Value[9])}))
  period = as.numeric(lapply(x, function(i){return(i$Value[10])}))
  
  #correct p-value from lsp by adding 1/perm
  p_correct = ( (ps*perm)+1 ) / (perm+1)
  
  print(paste0( length(which(p_correct < 0.01)), " unadjusted p-values < 0.01"))
  
  p_adjust = p.adjust(p_correct, method = "BH")
  print(paste0( length(which(p_adjust < 0.01)), " adjusted p-values < 0.01"))
  print(paste0( length(which(p_adjust < 0.005)), " adjusted p-values < 0.005"))
  
  return(cbind(p_adjust, power, period))
}

#
p1 = p_lsp_fun(rando_smooth_circ, perm=2000)
p2 = p_lsp_fun(rando_smooth_tidal, perm=2000)

#also read in the other rhythmicity tests implemented in BioDare2

R24 = read.csv("BioDare/BD2_eJTK.csv", skip=20)
R35 =  read.csv("BioDare/BD2_18_35.csv", skip=20)

MFF_28 = read.csv("BioDare/MFF_20_28.csv", skip=24)
MFF_34 = read.csv("BioDare/MFF_18_34.csv", skip=24)
FFT_28 = read.csv("BioDare/FFT_20_28.csv", skip=25)

metadata <- data.table( ID = df$ID,
                        BioDare_ID = 1:nrow(df),
                        Offset = str_split(df$X, pattern="_", n=2, simplify=TRUE)[,1],
                        Condition = str_split(df$X, pattern="_", n=2, simplify=TRUE)[,2], 
                        p_28 = p1[,1],
                        p_tidal = p2[,1],
                        power_28 = p1[,2],
                        power_tidal = p2[,2],
                        p_24_BD2 = R24$emp.p.BH.Corrected,
                        p_35_BD2 = R35$emp.p.BH.Corrected,
                        MFF_period_28 = MFF_28$Period,
                        MFF_period_34 = MFF_34$Period,
                        MFF_phase_28 = MFF_28$Circ..Phase.To.Zero,
                        MFF_phase_34 = MFF_34$Circ..Phase.To.Zero,
                        FFT_period = FFT_28$Period,
                        FFT_phase = FFT_28$Circ..Phase.To.Zero,
                        FFT_error = FFT_28$ERR,
                        key = "ID")

metadata$Group = paste(metadata$Offset, metadata$Condition, sep="_")

write.csv(metadata, "metadata_temperature.csv")
```

Read in the metadata. "Big" means a cycle from 8-32C, "small" means 14-26C. 
```{r}

metadata = read.csv("metadata_temperature.csv")
metadata = data.table(metadata, key="ID")

#rows 1-104 are individual animals, rows 105-108 are group means
no_mean = metadata[1:104,]
sig = no_mean[p_24_BD2 < 0.001,]

#note that groups are not the same size
groups = unique(no_mean$Group)

n_sig = sapply(groups, function(g){
  n = nrow(metadata[p_28 < 0.001 & Group == g, ])
  return(n)
})
n_sig

n_sig = sapply(groups, function(g){
  n = nrow(metadata[p_24_BD2 < 0.001 & Group == g, ])
  return(n)
})
n_sig

#not surprisingly, the requirement that periods are exactly 24h is more stringent.

ggplot(sig, aes(Condition, MFF_period_28, fill= Group)) + 
  geom_boxplot(outlier.colour = NA) +
  geom_jitter(alpha=.5) + theme_bw()

#Even among animals with periods at 24, there are a range of periods. This is partly due to noise and the short-ish nature of the time series, as well as true biological variability (certainly for free-running). 

ggplot(metadata, aes(Condition, power_28, fill= Group)) + 
  geom_boxplot(outlier.colour = NA) +
  geom_jitter(alpha=.5) + theme_bw()

ggplot(metadata, aes(Condition, FFT_error, fill= Group)) + 
  geom_boxplot(outlier.colour = NA) +
  geom_jitter(alpha=.5) + theme_bw()

wilcox.test(sig[Group=="Small_constant",]$power_28, sig[Group=="Small_cycle",]$power_28)
#p=0.076
wilcox.test(metadata[Group=="Big_constant",]$power_28, metadata[Group=="Big_cycle",]$power_28)
#p=0.017

```
Let's do some simple statistics.

```{r}
#convert phases to radians
conv <- 2*pi/24

circ.mean <- function (x) 
{
  sinr <- sum(sin(x))
  cosr <- sum(cos(x))
  circmean <- atan2(sinr, cosr)
  circmean
}

#calculate circular means of phases. 
circ_means <- sapply(unique(sig$Group), function(x){
  
  circ.mean(conv*(sig[Group == x,]$MFF_phase_28))/conv
})

#also, make a column that puts the phases into circular data format
metadata$circ = as.circular(metadata$MFF_phase_28, units = "hours", rotation="clock")
no_mean = metadata[1:104,]

sig=no_mean[p_24_BD2 < 0.001,]

#Rayleigh test for circular uniformity
sapply(groups, function(g){
  res = rayleigh.test(sig[Group==g]$circ)
  return(res$p.value)
})

#Significant for small cycle and marginally for big_cycle (our sample sizes are small). Nonsig. for free-running indicates that individuals with free-running rhythms are not synchronized with each other. We can visualize this:

data.segm <- data.frame(x=0,y=0, circ_phase = circ_means,yend=1.5, Group = names(circ_means), 
                        Offset = str_split(names(circ_means), pattern="_", n=2, simplify=TRUE)[,1],
                        Condition = str_split(names(circ_means), pattern="_", n=2, simplify=TRUE)[,2])

data.segm$Condition = factor(data.segm$Condition, levels=c("cycle", "constant"), labels=c("Cycle", "Free-run"))
data.segm$Offset= factor(data.segm$Offset, levels=c("Small", "Big"), labels=c("14-26°C", "8-32°C"))

sig$Condition = factor(sig$Condition, levels=c("cycle", "constant"), labels=c("Cycle", "Free-run"))
sig$Offset= factor(sig$Offset, levels=c("Small", "Big"), labels=c("14-26°C", "8-32°C"))

pl <- ggplot(sig, aes(x = MFF_phase_28, y=1, colour=Condition)) + geom_point(size=1.8) + theme_bw() + theme(axis.title.y=element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank())

pl <- pl + geom_segment(data=data.segm, aes(x=(circ_phase + 24) %% 24, y=0, xend=(circ_phase + 24) %% 24, yend=1.5), inherit.aes=FALSE)

png(res=300, width=4, height=8, units='in', file="F:/circadian/plots/Temperature_phases_p24_01.png")
pl + coord_polar() + scale_x_continuous("", breaks = c(0, 6, 12, 18), limits = c(0, 24)) + facet_grid(Offset~Condition) + theme(legend.position = "none")
dev.off() 

#We can also use circacompare to test for differences in phase between mean time series (using cosinors). need to fiddle with the formatting a bit. 

phase_df = data.frame(matrix(NA, ncol = 3, nrow = 4 * 78))
names(phase_df) = c("time", "measure", "group")
phase_df$time = 6:83

smooth <- read.csv("smoothed_temperature.csv")

just_means = smooth[grepl("mean", smooth$X.1),]
rownames(just_means) = just_means$X.1

phase_df$measure = as.vector( t( just_means[,3:80] ) )

phase_df$group = rep(rownames(just_means), each=78)

##compare the two temperature cycles... 
out <- circacompare(x = phase_df[phase_df$group == "Small_cycle_mean" | phase_df$group == "Big_cycle_mean", ], col_time = "time", col_group = "group", col_outcome = "measure", alpha_threshold = 0.3)

#both time series are highly rythmic according to this method as well.
#There is actually a highly significant difference between the phases, but just of 3h (ZT15 for big, ZT18 for small). 

out <- circacompare(x = phase_df[phase_df$group == "Small_cycle_mean" | phase_df$group == "Small_constant_mean", ], col_time = "time", col_group = "group", col_outcome = "measure", alpha_threshold = 0.3)

out
#circacompare doesn't think there is a free-running rhythm at 24 (p=0.14). If we force it to, it thinks there is a phase shift of +8h in free-run. Remember, the LSP approach (which is significant) tested for periods between 20-28h, not 24h exactly.

out <- circacompare(x = phase_df[phase_df$group == "Big_cycle_mean" | phase_df$group == "Big_constant_mean", ], col_time = "time", col_group = "group", col_outcome = "measure", alpha_threshold = 0.3)

#similar here, although even circacomapre thinks the free-running is rhythmic. This time the phase shift is just +5. 
out


```


And make some other figures...
```{r}

df = read.csv("activity_datasheet.csv")

names(df)[3] = "Raw_distance"
names(df)[4] = "Filtered_distance"
names(df)[7] = "Percent_active"
names(df)[9] = "Normalized_raw_movement"
names(df)[10] = "Normalized_filtered_movement"

smooth_fun = function(j){
  smooth = cma(y=as.numeric(j) , order=4)
  return(smooth$fitted)
}

#hooo boy 
df <- df %>% group_by(ID) %>%  mutate(smoothed = smooth_fun(Normalized_filtered_movement))
#df <- df %>% group_by(ID) %>% mutate(Det = detrend(smoothed))
df = data.table(df, key = "ID")

metadata$Condition = factor(metadata$Condition, levels=c("cycle", "constant"), labels=c("Cycle", "Free-run"))
metadata$Offset= factor(metadata$Offset, levels=c("Small", "Big"), labels=c("14-26°C", "8-32°C"))


dt = behavr(df[df$ID %in% metadata$ID,], metadata)
dt$t = dt$bin * 3600

ggetho(dt, summary_time_window = hours(1), aes(x = t, y = interaction(ID, Offset, Condition, sep = " : "), z=smoothed)) +
  stat_tile_etho()

#to group, change Y axis
pl <- ggetho(dt, summary_time_window = hours(1), aes(x = t, y = interaction(Offset, Condition, sep = " : "), z=smoothed)) +
  stat_tile_etho() + stat_ld_annotations() + ggetho::scale_x_hours()

pl



animal = sig[sig$Group=="Small_cycle",]$ID[ which.max(sig[sig$Group=="Small_c",]$power_28) ]

png(file="F:/circadian/plots/Temp_small.png", res=150, units="in", width=24, height=36)
p <- ggetho(dt[dt$ID %in% sig[sig$Offset=="Small",]$ID,], summary_time_window = hours(1), aes(x = t, y = smoothed, colour=Condition)) +
  stat_pop_etho() + stat_ld_annotations() +
  facet_grid(ID ~ Condition) + ggetho::scale_x_hours() + theme_minimal() + theme(legend.position = "none")
p
dev.off()

png(file="F:/circadian/plots/Temperature_plot_smooth.png", res=300, units="in", width=8, height=8)
ggetho(dt, summary_time_window = hours(1), aes(x = t, y = smoothed, colour=Condition)) +
  stat_pop_etho() + stat_ld_annotations() +
  facet_grid(Offset ~ .) + ggetho::scale_x_hours() + theme_bw()
dev.off()


png(file="F:/circadian/plots/Temperature_bars_indiv.png", res=300, units="in", width=8, height=8)
ggetho(dt, summary_time_window = hours(1), aes(x = t, y = smoothed, colour=Condition)) +
  stat_ld_annotations(height=1, alpha=0.3, outline=NA) + stat_pop_etho() +
  facet_grid(Offset ~ Condition) + ggetho::scale_x_hours() + theme_bw()
dev.off()

#will colour in the rectangles with the temperature gradient later
png(file="F:/circadian/plots/Temperature_indiv.png", res=300, units="in", width=8, height=8)
ggetho(dt, summary_time_window = hours(1), aes(x = t, y = smoothed, colour=Condition)) +
  stat_ld_annotations(ld_colours = c("white", "white")) + stat_pop_etho() +
  facet_grid(Offset ~ Condition) + ggetho::scale_x_hours() + theme_bw()
dev.off()

png(file="F:/circadian/plots/Temperature_indivs_minimal.png", res=300, units="in", width=8, height=8)
ggetho(dt, summary_time_window = hours(1), aes(x = t, y = smoothed, colour=Condition)) +
  stat_pop_etho() +
  facet_grid(Offset ~ Condition) + ggetho::scale_x_hours() + theme_minimal()
dev.off()

png(file="F:/circadian/plots/Temperature_day_plot.png", res=300, units="in", width=4, height=4)
pl <- ggetho(dt, summary_time_window = hours(1), aes(x=t, y=smoothed, colour=Condition), time_wrap = hours(24)) +
  facet_grid(Offset ~ .) + stat_ld_annotations(height=1, alpha=0.3, outline=NA) + stat_pop_etho() + theme_bw()
pl
dev.off()

#for "sig" individuals...

sig = metadata[p_24_BD2 < 0.001,]
df_sig = df[df$ID %in% sig$ID,]
dt = behavr(df_sig, sig)
dt$t = dt$bin * 3600

png(file="F:/circadian/plots/Temperature_sig_indiv.png", res=300, units="in", width=8, height=8)
ggetho(dt, summary_time_window = hours(1), aes(x = t, y = smoothed, colour=Condition)) +
  stat_ld_annotations(ld_colours = c("white", "white")) + stat_pop_etho() +
  facet_grid(Offset ~ Condition) + ggetho::scale_x_hours() + theme_bw()
dev.off()


png(file="F:/circadian/plots/Temperature_day_plot_sig_split.png", res=300, units="in", width=4, height=4)
pl <- ggetho(dt, summary_time_window = hours(1), aes(x=t, y=smoothed, colour=Condition), time_wrap = hours(24)) +
  facet_grid(Offset ~ .) + stat_ld_annotations(height=1, alpha=0.3, outline=NA) + facet_grid(Offset ~ Condition) + stat_pop_etho() + theme_bw()
pl
dev.off()

```

```{r}
##while we're at it, let's briefly re-calculate the mean series from just the sig. individuals (we can see that the data are less noisy, which is expected)

sig_df <- read.csv("Sig_datasheet.csv")


##applies centered moving average of order 4
smoothed = apply(sig_df, 1, function(j){
  
  dat = j[-c(1,2,3)]
  
  NAs = is.na(dat)
  
  smooth = cma(y=as.numeric(t(dat[!(NAs)]) ) , order=4)
  
  dat[!(NAs)] = smooth$fitted
  
  return(dat)
})

colnames(smoothed) = sig_df$ID

#write.csv(t(smoothed), file="Sig_smoothed.csv")

#just means
smoothed <- smoothed[,grepl("mean", colnames(smoothed))]


rando_smooth_circ = apply(smoothed, 2, function(j, n=2000){
  tmp = randlsp(repeats = n, x=as.numeric(j), type="period", ofac=50, from = 20, to=28, plot=FALSE, trace=FALSE)
  return(summary(tmp))
})


# perm is number of permutations done above
p_lsp_fun <- function(x, perm=2000){
  ps = as.numeric(lapply(x, function(i){return(i$Value[13])}))
  power = as.numeric(lapply(x, function(i){return(i$Value[9])}))
  period = as.numeric(lapply(x, function(i){return(i$Value[10])}))
  
  #correct p-value from lsp by adding 1/500 (or whatever)
  p_correct = ( (ps*perm)+1 ) / (perm+1)
  
  print(paste0( length(which(p_correct < 0.01)), " unadjusted p-values < 0.01"))
  
  p_adjust = p.adjust(p_correct, method = "BH")
  print(paste0( length(which(p_adjust < 0.01)), " adjusted p-values < 0.01"))
  print(paste0( length(which(p_adjust < 0.005)), " adjusted p-values < 0.005"))
 
  return(cbind(p_adjust, power, period))
}


p1 = p_lsp_fun(rando_smooth_circ, perm=2000)




```
